Variance reduction for Markov chains via martingale representations

### 2020-05-02 aap/Revision/vr-mcmc-revision-new-new.tex
update

\begin{lem}
\label{lem:exp_strongly_convex}
Let $(X_p^x)_{p \geq 0}$ be the ULA-generated Markov chain with ... . Then for any $\lambda \in \rset$, it holds
\begin{equation}
\label{eq:exp_strongly_convex}
\PE_x \rme^{\lambda\left(\pi_n^N(f) - \PE_x\pi_n^N(f)\right)} \leq e^{\frac{4\lambda^2 \|f\|^2_{\text{Lip}}\left(1+\frac{N}{n}\right)d(m+L)^2}{\gamma n m^2L^2}}
\end{equation}
\end{lem}
\begin{proof} Note that for ULA algorithm the distribution $\P_x$ is Gaussian for any $x$, thus, following \cite[Proposition 3]{dm2016HighDim}, we may bound the Wasserstein-1 distance between $\P_x$ and $\P_y$ by
\[
W_1(\P_x,\P_y) = W_1(\mathcal{N}(x-\gamma\mu(x),\gamma),\mathcal{N}(y-\gamma\mu(y),\gamma)) = \|x-y-\gamma(\mu(x)-\mu(y))\| \leq \sqrt{1-\kappa}\|x-y\|
\]
where $\kappa = \frac{2mL\gamma}{m+L}$. Hence, it holds
\[
W_1(\P_x,\P_y) \leq \left(1-\frac{\kappa}{2}\right)\|x-y\|
\]
Thus, we might apply the result of \cite[Theorem 4]{joulin2010}, yielding for any $\lambda \in \rset$  
\[
\PE_x \rme^{\lambda\left(\pi_n^N(f) - \PE_x\pi_n^N(f)\right)} \leq e^{4\lambda^2 \|f\|^2_{\text{Lip}} V^2}
\]
with the quantity $V^2 = \frac{4}{\kappa n}\left(1 + \frac{N}{n}\right)\sup_{x}\frac{\sigma^2(x)}{n_x \kappa}$, $\sigma^2(x) = \frac{1}{2}\iint \|y-z\|^2P_x(dy)P_x(dz) = \gamma d$ and $n_x \geq 1$.
\end{proof}

@@

Next we use \cite[Proposition 1]{joulin2010}, stating that

>>>>

@@

Note that for ULA algorithm the distribution $\P_x$ is Gaussian for any $x \in \rset^d$, thus, following \cite[Proposition 3]{dm2016HighDim}, we may bound the Wasserstein-1 distance between $\P_x$ and $\P_y$ by
W_1(\P_x,\P_y) = W_1(\mathcal{N}(x-\gamma\mu(x),\gamma),\mathcal{N}(y-\gamma\mu(y),\gamma)) = \|x-y-\gamma(\mu(x)-\mu(y))\| \leq \sqrt{1-\kappa\gamma}\|x-y\|
where $\kappa = \frac{2mL}{m+L}$. Hence, it holds
W_1(\P_x,\P_y) \leq \left(1-\frac{\kappa\gamma}{2}\right)\|x-y\|
which means that the Markov kernel $\P$ is a strict contraction in Wasserstein metric. This allows to use \cite[Proposition 1]{joulin2010}, stating that
